---
title: "Course Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the data
Here we load the data into R

```{r data,results="hide"}
a<-read.csv("./activity.csv")
```

## The mean total number of steps taken per day

Calculate the total number of steps taken per day 

```{r sumsteps}
    
## calculate the total steps per day
asum<-tapply(a$steps,a$date,sum,na.rm=T)
asum

```
Show the histogram

```{r histogram}

## Show the histogram
hist(asum,main="",xlab="Total number of steps per day" )

```

Calculate the mean and median of the total number of steps taken per day


```{r meansteps}

## calculate the mean per day
amean<-tapply(a$steps,a$date,mean,na.rm=T)
amean
## calculate the median per day
amedian<-tapply(a$steps,a$date,median,na.rm=T)
amedian
```

## The average daily activity pattern



```{r plot}
as<-tapply(a[,1],as.factor(a[,3]),mean,na.rm=T)
plot(a[1:length(as), "interval"], as,
     type = "l",
     ylab = "Average number of steps",
     xlab = "Time interval")
```

The interval containing the maximum number of steps

```{r max}
a[which.max(as),3]
```


## Imputing missing values

Calculate the total number of missing values in the dataset 

```{r}
sum(is.na(a[,1]))
```

Filling in all of the missing values in the dataset.

```{r}
a2<-cbind(a)
for (i in 1:length(a2[,1])){
    if (is.na(a2[i,1])) {
        if ((i %% 288)==0){
            a2[i,1]<-as[288]
        } else {
             a2[i,1]<-as[i %% 288]
        }
        
    }
    
}
```

Make a histogram of the total number of steps taken each day

```{r}
a2sum<-tapply(a2$steps,a2$date,sum)
hist(a2sum,main="",xlab="Total number of steps per day" )
```


Calculate the mean and median of the total number of steps taken per day


```{r}

## calculate the mean per day
a2mean<-tapply(a2$steps,a2$date,mean)
a2mean
## calculate the median per day
a2median<-tapply(a2$steps,a2$date,median)
a2median
```
The histogram has nicer shape when we use the filled data.


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend"

```{r}
a2[,"date"]<-as.Date(a2[,"date"],format="%Y-%m-%d")
wd<-as.POSIXlt(a2$date)$wday
w<-vector(length=length(wd))
w[((wd<6)&(wd>0))]<-"weekday"
w[((wd<1)|(wd>5))]<-"weekend"
a2$day<-as.factor(w)
head(a2$day)
```
Make a panel plot containing a time series plot  of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.



```{r}
sp<-split(a2,a2$day)
p1<-tapply(sp[[1]]$steps,sp[[1]]$interval,mean)
p2<-tapply(sp[[2]]$steps,sp[[2]]$interval,mean)
par(mfrow=c(2,1))
plot(as.integer(names(p2)), p2,main="Weekend",
     type = "l",
     ylab = "Number of steps",
     xlab = "Interval")
plot(as.integer(names(p1)), p1,main="Weekday",
     type = "l",
     ylab = "Number of steps",
     xlab = "Interval")

```


